### 2021-12-01 deliverables/presentation.Rmd
Added bare templates for paper and presentation

---
title: "Time Series Bootstrapping For Solar Current"
author: "Evan Kessler, Adam Kiehl, and Boston Lee"
date: "2021/12/9"
output: 
  beamer_presentation: default
---

# Background