### 2021-12-05 deliverables/presentation.Rmd
Initial rough presentation content

---
title: "Time Series Bootstrapping For Solar Current"
author: "Evan Kessler, Adam Kiehl, and Boston Lee"
date: "2021/12/9"
output: 
  beamer_presentation: default
---

# Research Question

- SolarStat: Modeling Photovolatic Sources through Stochastic Markov Processes
    - Marco Miozzo, Davide Zordan, Paolo Dini, and Michele Rossi
- Their objective: Describe how solar irradiance is translated into current output in a PV system throughout the day and cluster results with a Markov model to determine day/night breaks.
- Our objective: Describe how solar irradiance is translated into current output in a PV system and bootsrap results both with and without respect to time, and compare.

# Background

- Understanding renewable energy sources like solar power is critical to the next century. 
- PV modules installed at individual households/nodes reduce the overall load on the electrical grid. 
- Balancing the expected contribution of an individual PV module with the expected residual demand required of the grid is a large optimization problem. 
    - We dealt with half of that optimization problem involving the expected contribution of the PV module.

# Data

- National Solar Radiation Database (NSRD) from National Renewable Energy Laboratory (NREL) accessed through API
- Los Angeles data (1998-2010) was used to replicate the source paper. (IS THE YEAR RANGE RIGHT BOSTON?)
    - Date and time, direct normal irradiance (DNI), and temperature data used.

# Preprocessing

INSERT FIG 1 FROM PAPER HERE
- Astronomical adjustments performed on DNI to find effective solar irradiance:
    - Latitude, longitude, time zone, and the tilt of the earth used in calculations. 
    - $I_{eff} \propto I_{sun}cos(\theta)$

# Photovolatic (PV) Module

- Power processing units are attached to PV modules to optimize power output for a given current input.
    - Maximizes power $P=IV$ over a range of possible voltages. 
- Result of interest is current output after being adjusted by the power processor and an assumed leakage rate.

# Bootstrapping

# Results

# Future Research

- A similar procedure could be performed on a different city's data like Fort Collins.
- A similar procedure could be performed over a different range of years. A comparison may reveal the effects of global warming. 
- Our original source paper examined multiple PV modules simultaneously in the context of a larger grid. This examines both sides of the contribution/demand optimization problem.