### 2021-12-08 deliverables/presentation.Rmd
Added notes on bootstrapping

---
title: "Time Series Bootstrapping For Solar Current"
author: "Evan Kessler, Adam Kiehl, and Boston Lee"
date: "2021/12/9"
output:
  beamer_presentation: default
bibliography: references.bib
---

# Research Question

- SolarStat: Modeling Photovolatic Sources through Stochastic Markov Processes
    - Marco Miozzo, Davide Zordan, Paolo Dini, and Michele Rossi
- Their objective: Describe how solar irradiance is translated into current output in a PV system throughout the day and cluster results with a Markov model to determine day/night breaks.
- Our objective: Describe how solar irradiance is translated into current output in a PV system and bootsrap results both with and without respect to time, and compare.

# Motivation

- Understanding renewable energy sources like solar power is critical to the next century.
- PV modules installed at individual households/nodes reduce the overall load on the electrical grid.
- Balancing the expected contribution of an individual PV module with the expected residual demand required of the grid is a large optimization problem.
    - We dealt with half of that optimization problem involving the expected contribution of the PV module.

# Background

# Data

- National Solar Radiation Database (NSRD) from National Renewable Energy Laboratory (NREL) accessed through API
- Los Angeles data (1998-2010) was used to replicate the source paper.
    - Date and time, direct normal irradiance (DNI), and temperature data used.

# Preprocessing

INSERT FIG 1 FROM PAPER HERE
- Astronomical adjustments performed on DNI to find effective solar irradiance:
    - Latitude, longitude, time zone, and the tilt of the earth used in calculations.
    - $I_{eff} \propto I_{sun}cos(\theta)$

# Photovolatic (PV) Module

- Power processing units are attached to PV modules to optimize power output for a given current input.
    - Maximizes power $P=IV$ over a range of possible voltages.
- Result of interest is current output after being adjusted by the power processor and an assumed leakage rate.

# Bootstrapping

- We chose our initial paper because of its use of Monte Carlo. After working with it for a little it was brought to our attention that Monte Carlo was used in conjunction with a backward/forward sweep method in order to calculate the pdf. After some confusion it was decided to switch to another paper dealing with the same data/motivation but without said methods (including Monte Carlo).
- Luckily, the week after we switched papers we begun learning about bootstrapping, a way to sample from our data while taking into consideration our time change.

INSERT GRAPHS FROM IID BOOTSTRAPPING AND BLOCK HERE
ALSO INSERT FIG 1 FROM PAPER

- The paper focused most of their sampling on night-day clustering, showing the difference between night and day through current. Night yielded thousands of 0 current values, leading to skewed data when sampling. Instead we took too approaches to bootstrap the current vs. hour data.
- One using non-overlapping blocks sampling from the 8 hours surrounding the peak irradiance. The other being IID, yielding dependence in our data.

# Results

# Future Research

- A similar procedure could be performed on a different city's data like Fort Collins.
- A similar procedure could be performed over a different range of years. A comparison may reveal the effects of global warming.
- Our original source paper examined multiple PV modules simultaneously in the context of a larger grid. This examines both sides of the contribution/demand optimization problem.